---
title: "Tornado Data EDA"
author: "JMG"
format: html
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(ggthemes)
library(skimr)
library(maps)

theme_set(theme_minimal(base_size = 12))
```

We obtained the raw data from [here](https://www.spc.noaa.gov/wcm/data/1950-2022_actual_tornadoes.csv)

## The Data

Let's load the data:

```{r}
tornados <- read_csv("data/tornados.csv")
```

Check what we read in

```{r}
glimpse(tornados)
```

This data contains 68,693 rows and 27 columns (variables).

```{r}
skim(tornados)
```

Use Console when you use the skim command. It can get ugly real quick.

```{r}
tornados %>%
  ggplot(aes(x = st)) +
  geom_bar() +
  coord_flip()
```

Makes a bar plot of the count of tornados by count. Virgin Islands, Washington DC and Cuban Islands included.

```{r}
tornados %>%
  ggplot(aes(x=slon,y=slat)) +
  geom_point(alpha=0.2)
```

Makes a really cool map

```{r}
states_df <- map_data("state")
```

Making the states

```{r}
tornados49 <- tornados %>%
  filter(st != "AK", st != "HI", st != "PR", st != "VI")
```

Apparently we don't like Alaska, Puerto Rico, the Virgin Islands, and the HI anymore.

```{r}
tornados49 %>%
  ggplot(aes(x = st)) +
  geom_bar() +
  coord_flip()
```

Bar plot barring the four states 

```{r}
tornados49 %>%
  group_by(st) %>%
  summarise(per_state = n()) %>%
  arrange(desc(per_state))
```

Putting the bar graph into numbers and arranging

```{r}
tornados49 %>%
  group_by(st) %>%
  summarise(per_state = n()) %>%
  arrange(desc(per_state)) %>%
  ggplot(aes(x=reorder(st,per_state),y=per_state)) +
  geom_bar(stat = "identity")  +
  coord_flip()
```

```{r}
tornados_49 <- tornados49 %>%
  filter(slon >= -105)
states_of <- states_df %>%
  filter(long >= -105)
glimpse(states_of)
```

```{r}
tornados_49 %>%
  group_by(st) %>%
  summarise(per_state = n()) %>%
  arrange(desc(per_state)) %>%
  ggplot(aes(x=reorder(st,per_state),y=per_state)) +
  geom_bar(stat = "identity")  +
  coord_flip()
```

```{r}
tornados %>%
  filter(fat > 0) %>%
  ggplot(aes(x=fat)) +
  geom_bar()
```

```{r}
tornados49 %>%
  mutate(mag = factor(mag)) %>%
  ggplot(aes(x=mag,fill=fc)) +
  geom_bar(position = position_dodge(preserve = "single")) +
  scale_y_log10()
```

```{r}
tornados %>%
  mutate(ns = factor(ns)) %>%
  ggplot(aes(x=ns)) +
  geom_bar() +
  scale_y_log10()
```

```{r}
tornados_49 %>%
  ggplot(aes(x=len)) +
  geom_histogram() +
  scale_x_log10()
```

```{r}
tornados49 %>%
  mutate(mag=factor(mag)) %>%
  ggplot(aes(x=mag,y=len)) +
  geom_boxplot()
```

```{r}
tornados49 %>%
  mutate(mag=factor(mag)) %>%
  ggplot(aes(x=mag,y=fat)) +
  geom_point()
```

